---
title: "BGGN213_Class8"
author: "Yutao Wen"
date: "4/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Clustering - Kmeans Function

```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```

```{r}
kmeans(x, centers=2, nstart=20)
km <- kmeans(x, centers=2, nstart=20)

```


> Q. How many points are in each cluster?

> Q. What ‘component’ of your result object details
      - cluster size? 
      
```{r}
km["size"]

```
     
      - cluster assignment/membership? 
      
```{r}
km["cluster"]
```

      - cluster center? 
      
```{r}
km$centers
```

> Plot x colored by the kmeans cluster assignment and
      add cluster centers as blue points
      
      
```{r}
# Plot scatterplot colored by different clusters
plot(x, col=km$cluster)

# Add center points to the plot
points(km$centers, col="blue", pch=18, cex=3)

```



## Hierarchical Clustering - hclust Function

Here we dont have to spell out K, the number of clusters but we need to give it a distance matrix as input

```{r}
d <-  dist(x)
hc <- hclust(d)
plot(hc)
```

```{r}
# Draw a line at height = 6 and cut the tree under
plot(hc)
abline(h=6, col="red")
cutree(hc, h=6)
# or can use plot(hc, k=2)
```


```{r}
gp2 <- cutree(hc, k=2)
gp2
gp3 <- cutree(hc, k=3)
gp3
```

```{r}
table(gp2)
table(gp3)
table(gp2, gp3)
```

Practice with another more complicated sample

```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
  matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2),   # c1
  matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
  matrix(c(rnorm(50, mean = 1, sd = 0.3),           # c3
           rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
plot(x)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```

```{r}
dis <- dist(x)
hcl <- hclust(dis)
plot(hcl)
abline(h=4)
ct2 <- cutree(hcl, k=2)
ct3 <- cutree(hcl, k=3)
table(ct2, ct3)
plot(x, col=ct3)

```


## PCA - prcomp function

```{r}
mydata <- read.csv("https://bioboot.github.io/bggn213_S19/class-material/expression.csv", row.names = 1)
head(mydata)

# transpose the data
head(t(mydata))
```

There are `r nrow(mydata)` genes in my dataset

```{r}
pca <- prcomp(t(mydata), scale=TRUE)
summary(pca)
```

```{r}
attributes(pca)
```

Let's make our first PCA plot

```{r}
plot(pca$x[,1], pca$x[,2])
# Variance captured per PC
 pca.var <- pca$sdev^2
 pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
 pca.var.per
```

```{r}
xlab <- paste("PC1(", pca.var.per[1],")%")
ylab <- paste("PC2(", pca.var.per[2],")%")
plot(pca$x[,1], pca$x[,2], xlab = xlab, y = ylab, col=mycols)
```

## UK Food Practice

```{r}
x <- read.csv("https://bioboot.github.io/bggn213_S19/class-material/UK_foods.csv", row.names = 1)
head(x)
```

```{r}
dim(x)
```

```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```

```{r}
barplot(as.matrix(x), beside=F, col=rainbow(nrow(x)))
```

```{r}
pairs(x, col=rainbow(10), pch=16)
```

```{r}
PCA <- prcomp(t(x))
summary(PCA)
```

```{r}
# Plot PC1 vs PC2
plot(PCA$x[,1], PCA$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(PCA$x[,1], PCA$x[,2], colnames(x), col = c("yellow", "red", "blue", "green"))
```

```{r}
v <- round(PCA$sdev^2/sum(PCA$sdev^2)*100)
v
```

```{r}
z <- summary(PCA)
z$importance
```

```{r}
barplot(v, xlab="Principal Component", ylab="Percent Variation")
```

```{r}
PCA$rotation
```

```{r}
par(mar=c(10,3, 0.35, 0))
barplot(PCA$rotation[,1], las=2)
```

```{r}
par(mar=c(10,3, 0.35, 0))
barplot(PCA$rotation[,2], las=2)
```


```{r}
biplot(PCA)
```

