---
title: "BGGN213_Class7"
author: "Yutao Wen"
date: "4/24/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rescale <- function(x, na.rm=TRUE, plot=FALSE, ...) {
   rng <-range(x, na.rm=na.rm)
   answer <- (x - rng[1]) / (rng[2] - rng[1])
   if(plot) {
     plot(answer, ...)
}
   return(answer)
}
```

```{r}
rescale(1:10)
```

```{r}
#rescale(c(1,10,"string"))
```

```{r}
rescale2 <- function(x, na.rm=TRUE, plot=FALSE, ...) {
   if( !is.numeric(x) ) {
      stop("Input x should be numeric", call.=FALSE)
   }
   rng <-range(x, na.rm=na.rm)
   answer <- (x - rng[1]) / (rng[2] - rng[1])
   if(plot) {
      plot(answer, ...)
}
   return(answer)
}
```

```{r}
#rescale2(c(1,10,"string"))
```

## Function Practice

Write a function to identify NA elements in two vectors.

Start with a simple example input where I know what the answer is:

```{r}
x <- c(1,2,NA,3,NA)
y <- c(NA,3,NA,3,4)
```

```{r}
is.na(x)
is.na(y)
```

```{r}
is.na(x) & is.na(y)
```

Take sum of two to find how many of TRUE I have:

```{r}
sum(is.na(x) & is.na(y))
```

Create a function based on previous work:

```{r}
both_na <- function(x,y) {
  sum(is.na(x) & is.na(y))
}
```

```{r}
both_na(c(NA,NA,NA), c(NA,NA,NA))
```

Test abnormal cases of different lengths of vectors:

```{r}
both_na(c(NA,NA,NA), c(1,NA,NA,NA,NA,NA))
```
The shorter vector was recylcled to match with the longer one.

Check the lengths of the two vectors beforehand:

```{r}
x <- c(NA,NA,NA)
y <- c(1,NA,NA,NA,NA,NA)
length(x) != length(y)
```

Add that to the function:

```{r}
both_na2 <- function(x,y) {
  if(length(x) != length(y)) {
    stop("Input x and y should be the same length")
  }
  sum(is.na(x) & is.na(y))
}
```

```{r}
#both_na2(x,y)
```

Try the both_na3 function with extra features:
```{r}
both_na3 <- function(x, y) {
if(length(x) != length(y)) {
stop("Input x and y should be vectors of the same length")

}
  na.in.both <- ( is.na(x) & is.na(y) )
  na.number  <- sum(na.in.both)
  na.which   <- which(na.in.both)
  message("Found ", na.number, " NA's at position(s):",
          paste(na.which, collapse=", ") )
  return( list(number=na.number, which=na.which) )
}
```

```{r}
#both_na3(x,y)
```

```{r}
X <- c( 1, 2, NA, 3, NA) 
Y <- c(NA,3,NA,3, 4)
both_na3(X,Y)
```

Write a function grade() to determine an overall grade from a vector of student homework assignment scores dropping the lowest single alignment score.

```{r}
# Student 1
s1 <- c(100,100,100,100,100,100,100,90)
# Student 2
s2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
```

```{r}
grade <- function(x) {
  avg <- (sum(x, na.rm = TRUE)-min(x, na.rm = TRUE)) / (length(x)-1)
  return((avg))
}
```

```{r}
grade(s1)
grade(s2)
grade(c(s1,s2))
```

```{r}
url <- read.csv("https://tinyurl.com/gradeinput", row.names = 1)
ans <- apply(url, 1,grade)
sorted <- sort(ans, decreasing = TRUE)
sorted[1:5]
```

```{r}
# Start with a simple version of the problem
df1 <- data.frame(IDs=c("gene1", "gene2", "gene3"),
                  exp=c(2,1,1),
                  stringsAsFactors=FALSE)
df2 <- data.frame(IDs=c("gene2", "gene4", "gene3", "gene5"),
                  exp=c(-2, NA, 1, 2),
                  stringsAsFactors=FALSE)

# Simplify further to single vectors
x <- df1$IDs
y <- df2$IDs

intersect(x,y)

x%in%y

x[x%in%y]

cbind( x[ x %in% y ], y[ y %in% x ] )
```

```{r}
gene_interest <- function(x,y) {
  cbind( x[ x %in% y ], y[ y %in% x ] )
}
```

```{r}
merge(df1, df2, by="IDs")
```

```{r}
if (!requireNamespace("BiocManager"))
    install.packages("BiocManager")
BiocManager::install()
```

