---
title: '190419 Class6: R Functions'
author: "Yutao Wen"
date: "4/19/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Today we will focus on **R functions** but we will start with a bit of **file reading**.
```{r}
plot(1:10, type="l", col="blue")
```

```{r}
read.table("test1.txt", header = TRUE, sep=",")
```

Another file opening
```{r}
read.table("test2.txt", header = TRUE, sep = "$")
```

```{r}
read.table("test3.txt")
```

```{r}
read.csv("https://bioboot.github.io/bggn213_S19/class-material/test1.txt")
```

## Our first function

Add some numbers

```{r}
add <- function(x, y=1) {
    # The body
   x + y
}
  
```

```{r}
add(x=4, y=5)
```


```{r}
add(4)
```

```{r}
add(c(1,3,5), 1)
```

```{r}
#add(1,3,5)
```

```{r}
#add(x=1 y="yutao")
```

```{r}
rescale <- function(x) {
   rng <-range(x)
   (x - rng[1]) / (rng[2] - rng[1])
}
```

```{r}
rescale(1:10)
```

```{r}
rescale(c(1,3,NA,5,10))
```

```{r}
x <- c(1,3,NA,5,10)
rng <-range(x, na.rm = TRUE)
rng
(x - rng[1]) / (rng[2] - rng[1])
```

```{r}
rescale2 <- function(x, na.rm=TRUE) {
   rng <-range(x, na.rm = na.rm)
   (x - rng[1]) / (rng[2] - rng[1])
}
```

```{r}
rescale2(c(1,3,NA,5,10))
```

```{r}
rescale2(c(1,3,NA,5,10), na.rm = FALSE)
```

```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
   rng <-range(x, na.rm = na.rm)
   print("Hello")
   answer <- (x - rng[1]) / (rng[2] - rng[1])
   print("is it me what you looking for")
   if(plot) {
      plot(answer, typ="b", lwd=4)
     print("Don't sing please")
     
   }
   print("I can see it in ...")
   return(answer)
}
```

```{r}
rescale3(1:10)
```


```{r}
rescale3(1:10, plot = TRUE)
```

```{r}
rescale4 <- function(x, na.rm=TRUE, plot=FALSE) {
   rng <-range(x, na.rm = na.rm)
   print("Hello")
   answer <- (x - rng[1]) / (rng[2] - rng[1])
    return(answer)
   print("is it me what you looking for")
   if(plot) {
      plot(answer, typ="b", lwd=4)
     print("Don't sing please")
     
   }
   print("I can see it in ...")

}
```

```{r}
rescale4(1:10)
```

# Can you improve this analysis code?
library(bio3d)
s1 <- read.pdb("4AKE")  # kinase with drug
s2 <- read.pdb("1AKE")  # kinase no drug
s3 <- read.pdb("1E4Y")  # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")

Q1. What type of object is returned from the read.pdb() function? Q2. What does the trim.pdb() function do?
Q3. What input parameter would turn off the marginal black and grey rectangles in the plots and what do they represent in this case?
Q4. What would be a better plot to compare across the different proteins?
Q5. Which proteins are more similar to each other in their B-factor trends. How could you quantify this? HINT: try the rbind(), dist() and hclust() functions together with a resulting dendrogram plot. Look up the documentation to see what each of these functions does.

```{r}
library(bio3d)
s1 <- read.pdb("4AKE")  # kinase with drug
s2 <- read.pdb("1AKE")  # kinase no drug
s3 <- read.pdb("1E4Y")  # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

```{r}
is.vector(s1)
is.data.frame(s1)
is.list(s1)
```

```{r}
s1$seqres
```


```{r}
bfactor <- function(x) {
  library(bio3d)
  pdb <- read.pdb(x)
chainA <- trim.pdb(pdb, chain="A", elety="CA")
chainA.b <- chainA$atom$b
plotb3(chainA.b, sse=chainA, typ="l", ylab="Bfactor")
}
```

```{r}
bfactor("4AKE")
```



